Data Archiving Attitudes

### 2018-03-21 manuscript.Rmd
updates

---
title: "Data Archiving Attitudes"
author: "Heather A. Piwowar, Todd J. Vision"
date: "March 20, 2018"
output:
  md_document:
    variant: markdown_github
  html_document:
    css: style.css
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
csl: f1000-research.csl
bibliography: input_references.bib
---

@@

@@

The promise of data sharing and the role of data sharing policies

@@

@@

#### Dimensions of impact

A journal policy that requires authors to share the datasets that support their results as a condition of publication may impact the practices, attitudes, and opinions of data producing authors in many ways.  Here, we emphasize the short-term impacts on authors who publish in the policy-adopting journals.

The most direct impact is likely a change in frequency of data sharing, and a shift to data archiving locations specified in the policy.  Early studies suggest that authors who publish in journals with data sharing policies are more likely to share their data than authors who publish similar studies in similar journals `r citep(c(bib["piwowar2008a-revie"], bib["Piwowar_2011"]))`, but these investigations were correlative and did not attempt to directly measure the impact of journal policies over time.  Previous work emphasizes the importance of sharing data through best-practice archiving methods and locations.  Authors who share data upon request often discriminate against use and users `r citep(c(bib["campbell2008what-ar"], bib["reidpath2001data-sh"]))`.  The permanence of data sharing mechanisms is also a concern; email and lab-website based sharing choices have been shown to be impermanent solutions at best `r citep(c(bib["wren2008url-dec"], bib["wren2006e-mail"]))`.

@@

Upcoming policy change is a useful opportunity for policy evaluation

Several journals plan to adopt the Joint Data Archiving Policy in early 2011, as described at http://datadryad.org/jdap.  Participating journals include:

- American Naturalist
- Evolution
- Evolutionary Applications
- Heredity
- Journal of Evolutionary Biology
- Molecular Biology and Evolution.
- Molecular Ecology

@@

Rigorous research is needed into the impact of journal policies

@@

Research questions

@@

- How do authors’ attitudes, experiences, and practices around public data archiving change when the journals they publish in adopt mandatory data archiving policies?
- Are changes specific to the journals that implement the policies, or do they extend to other journals in the same subfield?

## Primary analysis
The primary analysis will be a seven-point Likert scale asking how strongly authors believe:  “It is the community norm in my field to publicly share datasets upon study publication by archiving all datasets online, for use by anyone for any purpose.”

About twelve other survey questions will allow secondary and exploratory analyses.

## Methods
Participants will be recruited from corresponding authors who publish in specific journals between November 2010 and November 2013.  Specifically, email addresses will be collected for papers published the previous month (or 3 previous months, for the initial month of recruitment) in each of the target journals through the ISI Web of Science.  No other identifying information (author names, article title, etc) will be retained.

Recruitment email lists will be compared across months, and emails only sent to those who have not previously received a recruitment email to reduce survey fatigue and ill-will.

The email invitation will include a URL where subjects can learn more about the study and then read a fact sheet, consent, and participate if they choose.  This email will also contain an “opt out” option, which will remove their address from the reminder list.  Those who have not “opted out” will be contacted by email again, a week later, politely asking them to take the survey if they haven’t already.

Individual participants can expect to spend 7-10 minutes maximum answering the online questionnaire.

Before the initial round of invitations is sent, I will inform journal editors of this ongoing survey and provide contact information in case they have questions, suggestions, or concerns.

Previous studies suggest 10%-30% of authors will complete the survey `r citep(c(bib["haeusslerinforma"], bib["wren2007the-wri"]))`.

Gender, ethnicity, and race are irrelevant to the study. Other than the requirement that participants be 18 years of age or older, age is irrelevant to the study. Medical history and/or conditions are irrelevant to the study.

To facilitate legal consent, this study is restricted to participants who are 18 years of age or older. The study is intended to establish knowledge about the implications of policies on active researchers.  No other criteria are necessary for inclusion; no criteria are grounds for exclusion.

## Methods 2

We expect the JDAP policies to be implemented in or about January 2011, though it will take several months to see any effect because the policies will likely be applied to newly submitted papers it will take multiple months for these papers to be published.  This timeline is likely to vary across JDAP-adopting journals depending on the length of their publication cycle.

Participants will be recruited from corresponding authors who publish in specific journals between November 2010 and November 2013.  An invitation to the survey will be issued once a month to all corresponding authors who published in specific journals the prior month. The responses of authors who published in JDAP-adopting journals will be compared, over time, to a) the responses of authors who publish in in-community journals that do not adopt JDAP, and b) the responses of authors who publish in outside-community journals that do not adopt JDAP.

### Data collection

The instrument used to collect data is be an online questionnaire. The questionnaire is designed to elicit the respondent’s practices, experiences, and opinions around public archiving of research datasets by asking about their opinion of community norms, recent behaviour with respect to a recently published paper, previous experiences with data sharing, and attitudes toward hypothetical situations.

@@

Aggregate answers will be compared over time, to understand if answers of authors in journals that adopt JDAP begin to differ from those of authors in comparable within-community and outside-community control journals.

The online survey will be administered through ViewsFlash® 6 by Cogix, hosted by Duke University.  There will be no way link the responses with the identities of participants.

### Limitations

The limitations of this study design include:

- There is limited time before JDAP is implemented (~January 2011), so I will only have two pre-JDAP-implementation control time points.  This limitation will be mitigated by the fact that there will be a lag between policy adoption is applicable to submitting authors and when their papers published.
- There are many variables that are also important to views of community norms and decisions about sharing that I am not collecting to keep the survey short and maintain anonymity. These include university characteristics, industry affiliations, subfield, publication rate, and age.
- This study makes no attempt to understand the rigor with which data sharing is encouraged in journals beyond their written policies:  many journals without policies nonetheless expect data sharing and many with written policies fail to enforce them.
- The opinions of corresponding authors with respect to data sharing may not be typical of all authors:  a future study could be done to quantify this.
- By only inviting authors to participate the first time they are listed as corresponding authors within our study scope, the population we are polling will shift gradually over time away from highly-prolific authors.
- Respondents may be reluctant to self-report when they do not share their datasets, particularly as their sense of sharing as a community norm increases, since survey respondents are often reluctant to report engaging in socially undesirable behavior.
- It will be difficult or impossible to derive causality direction for an association between reported data sharing behaviour and reported perception of data sharing as a community norm.
- Finally, this study only facilitates a relatively short-term look at policy effect:  it will likely take many more years for the effect of these policies to be fully felt.

Analysis of policy impact will be done using a modification of the commonly accepted difference-in-difference approach, as described by `r citep(bib["abadie2010synthet"])`.  The responses of authors who published in JDAP-adopting journals will be compared, over time, to a) the responses of authors who publish in in-community journals that do not adopt JDAP, and b) the responses of authors who publish in outside-community journals that do not adopt JDAP.  Because it is not clear which in-community and outside-community journals would provide the best counterfactual control of the JDAP-adopting journals, we will construct a “synthetic control” from a combination of responses across a subset of the control journals, as described in `r citep(bib["abadie2010synthet"])`.

### Sample size

The proposed study is subject to two different types of uncertainty:  uncertainty about the estimation of aggregate parameters, and also “uncertainty about the ability of the control group to reproduce the counterfactual of how the treated unit would have evolved in the absence of the treatment” (`r citep(bib["abadie2010synthet"])`. It is difficult to know how to conduct a formal power analysis in these conditions.  We have based our sample size decisions on the following information.

For populating the “treatment arm,” we plan to ask all corresponding authors of all JDAP-adopting journals.  The average number of papers published per month in the JDAP journals is 168.  Previous studies suggest 10%-30% of authors will complete the survey `r citep(c(bib["haeusslerinforma"], bib["wren2007the-wri"]))`.  If we assume a response rate of 20%, we will receive about 33 responses from the intervention group every month.  This would result in 95% confidence intervals around a Likert mean of 3.5 out of 7 ranges from 2.4 to 4.6.  If these confidence intervals prove too wide to enable effective comparisons, we could aggregate the monthly data into less frequent samples across time.

For populating the “control arms,” we recognize that the uncertainty is predominantly determined by the applicability of the controls.  Little information is currently available to determine, a priori, which journals will serve as appropriate controls.  As such, we suggest that surveying twice as many control journals as treatment journals may be an appropriate rule of thumb.  This balances the need to avoid overpowering the study (and thereby increasing survey-fatigue and effort) with a desire to ensure the most useful results (anticipating that unexpected future events may render some controls inappropriate during the three years of this study, as well as recognizing that the quality of inference improves as the number of available comparison units increases `r citep(bib["abadie2010synthet"])`.

Control journals will be selected by manual selection from a set of journals identified through ISI Journal Citation report categories as well as co-citation lists.  Attempts will be made to choose journals with comparable impact factors and inclination to produce quantitative datasets.

A 20% response rate and two control groups that are each twice as large as our treatment arm means we would receive about 33 treatment responses, 66 in-community control responses, and 66 outside-community responses per month.  In total, we expect about 165 responses per month.  Over the three years of data collection, this would lead to almost 6000 responses.  The actual number of responses will be less than this, because will only invite each corresponding author once.  On the other hand, the number of articles published by journals may increase over time and thus lead to a larger-than-anticipated sample.

Finally, we note that to the best of our knowledge, a similar study has not been administered and/or published.  If this sample size proves to be inadequate, the results can be viewed as pilot information to help inform future studies.

### Data sharing plan

Recruitment email addresses will be associated with the journal and month of publication.   No other identifying information (author names, article title, etc) will be retained from ISI Web of Science.   The researcher will not keep a record of participants’ names nor interface with them personally or by phone.  Questionnaire responses will be submitted via online web survey tool ViewsFlash®:  the recruitment email addresses will not be linked to their corresponding survey response, but only linked in aggregate (in groups defined by journal and month of publication).

Email addresses that are voluntarily provided by participants will be kept only for the duration of the prize drawing each month, and will not be linked to the survey responses in any way.

Survey responses will be associated with journal and month of publication.  The survey also asks the subjects to self-identify their job title (student, postdoc, tenured faculty, etc.)  We believe these datapoints will not be sufficient to permit reidentification, and will make no attempts to identify the respondents.

To maximize the usefulness of this research, we plan to make the anonymous questionnaire responses openly available in a data repository for reuse, after aggregating or obfuscating any responses that could otherwise facilitate identification.

### Recruitment

Participants will be recruited from corresponding authors who publish in specific journals between November 2010 and February 2012.  Specifically, email addresses will be collected for original research papers published the previous month (or 3 previous months, for the initial month of recruitment) in each of the target journals, using the ISI Web of Science.  The email addresses will be associated with the journal and month of publication.  No other identifying information (author names, article title, etc) will be retained.

Recruitment email lists will be compared across months, and emails only sent to those who have not previously received a recruitment email to reduce survey fatigue and ill-will.

A recruitment email will be sent to all identified email addresses, excluding those who had received a recruitment email in a prior month.  The email invitation will include a URL where subjects can learn more about the study and then read a fact sheet, consent, and participate if they choose.  This email will also contain an “opt out” option, which will remove their address from the reminder list.  Those who have not “opted out” will be contacted by email again, a week later, politely asking them to take the survey if they haven’t already.

Before the initial round of invitations is sent, we will inform journal editors of this ongoing survey and provide contact information in case they have questions, suggestions, or concerns.

Individual participants can expect to spend 7-10 minutes maximum answering the online questionnaire.

Previous studies suggest 10%-30% of authors will complete the survey `r citep(c(bib["haeusslerinforma"], bib["wren2007the-wri"]))`.

## Methods

# Introduction
Some researchers are sceptical regarding data archiving. We conducted a survey in order to find out how the adoption of data archiving policies by journals affects researcher's attitudes towards archiving and sharing data.

# The Survey

## Method

```{r, cache=TRUE, message=FALSE, echo=FALSE}

@@

The survey was carried out using [Survey Monkey](www.surveymonkey.com). Invitations to participate were sent out monthly between February 2011 and March 2014 to a selection of between $`r min(n_invitations_by_sent_month$N)`$ and $`r max(n_invitations_by_sent_month$N)`$ (with an average of $`r round(mean(n_invitations_by_sent_month$N))`$) authors of recently published articles in a pre-selected sample of journals. 
<!-- TODO: how were the journals selected -->
The invitations were sent out to a total of $`r nrow(invitation_timing)`$ authors, and $`r nrow(d)`$ responded (corresponding to a non-response rate of $`r round(100*nrow(d)/nrow(invitation_timing))`$\%).
Figure 1 shows the numbers of invitations and responses by the date of the author's last publication.
Participants filled in a questionnaire consisting of 15 multiple-choice questions regarding their attitudes, concerns and experiences with data archiving and sharing.

## Results

In the following, I will present an overview of the responses to questions 1, 2, and 9. The questions asked about the strength of agreement/disagreement with the following statements:

* _Question 1_ It is the community norm in my field to share, with qualified researchers, datasets that support the results of peer-reviewed research articles. Include data sharing that occurs by any mechanism (e.g. emailing datasets upon request, posting datasets to lab websites, including datasets in journal supplementary information or data repositories).
* _Question 2_ It is the community norm in my field to publicly archive all supporting datasets online, for use by anyone for any purpose, upon publication of a peer-reviewed research article.
* _Question 9_ I would seek alternative journals if JDAP was introduced.

@@

library(dplyr)
library(ggplot2)
library(reshape2)
source("./src/aux_functions.R")
source("./src/plotLikert.R")

responses <- c("Strongly Agree","Agree","Somewhat Agree","Somewhat Disagree","Disagree","Strongly Disagree") # response categories from top to bottom
question_labels <- c('Q_1'='"Data sharing is a community norm."\n(Question 1)',
                     'Q_2'='"Data archiving is a community norm."\n(Question 2)',
                     'Q_9'='"I would seek alternative journals\nif JDAP was introduced."\n(Question 9)')
tmp <- melt(d, measure.vars=c('Q_1','Q_2','Q_9')) #,'Q_12','Q_13'))
tmp$variable <- as.character(tmp$variable)
tmp$variable <- ordered(nmap(tmp$variable, question_labels), levels=question_labels)
#tmp$variable <- gsub("Q_", "Question ", tmp$variable) # gsub("Q_", "Question ", tmp$variable)

### Inferential statistics

```{r, message=FALSE, eval=TRUE, echo=FALSE}

@@

The following table shows the estimated coefficients and 95\%-confidence intervals (transformed to odds ratios) for responses to *question 2*. According to the model estimates, the introduction of the JDAP data archiving policy significantly increases the odds of a positive response regarding *data archiving* by a factor of $1.57$ (CI=$[1.33; 1.90]$), while the remaining two predictors do not show any statistically significant effects.

@@

# Appendix

## Appendix A: Questionnaire Structure
<!-- TODO: Add response options below. -->

How strongly do you disagree/agree with the following statements? (Regarding the next two questions.)

<!-- Field: Q4 -->
1. It is the community norm in my field to share, with qualified researchers, datasets that support the results of peer-reviewed research articles. Include data sharing that occurs by any mechanism (e.g. emailing datasets upon request, posting datasets to lab websites, including datasets in journal supplementary information or data repositories).

<!--
_(Strongly Agree - Agree - Somewhat Agree - Neutral - Somewhat Disagree - Disagree - Strongly Disagree)_
-->

responses_agree_disagree <- c("Strongly Agree","Agree","Somewhat Agree","Somewhat Disagree","Disagree","Strongly Disagree") # response categories from top to bottom

tmp <- melt(d, measure.vars=c('Q_1'))
tmp$variable <- gsub("Q_", "Question ", tmp$variable)
p <- plot_likert(tmp, response='value', var_x='label_PolicyAdoptedAtPublication', var_facet="variable", response_order=responses_agree_disagree, above_mid="Somewhat Agree", na.label="No response", xlab="% Responses", ylab="JDAP Policy")

<!-- Field: Q5 -->
2. It is the community norm in my field to publicly archive all supporting datasets online, for use by anyone for any purpose, upon publication of a peer-reviewed research article.

<!--
_(Strongly Agree - Agree - Somewhat Agree - Neutral - Somewhat Disagree - Disagree - Strongly Disagree)_
-->

responses_agree_disagree <- c("Strongly Agree","Agree","Somewhat Agree","Somewhat Disagree","Disagree","Strongly Disagree") # response categories from top to bottom

tmp <- melt(d, measure.vars=c('Q_2'))
tmp$variable <- gsub("Q_", "Question ", tmp$variable)
p <- plot_likert(tmp, response='value', var_x='label_PolicyAdoptedAtPublication', var_facet="variable", response_order=responses_agree_disagree, above_mid="Somewhat Agree", na.label="No response", xlab="% Responses", ylab="JDAP Policy")

<!-- Field: Q7 -->
3. Do any of your published research papers have publicly archived datasets?
  a. Yes, datasets from <b>5 or more papers</b> are publicly available on the internet
  b. Yes, datasets from <b>2-4 papers</b> are publicly available on the internet
  c. Yes, a dataset from <b>one paper</b> is publicly available on the internet  
  d. No, none of my paper-related datasets are publicly available on the internet
  e. I've never published a paper based on a dataset collected by me or my co-authors
  f. I'm not sure if any of my papers have publicly available datasets

<!-- Field: Q9 -->
4. To your knowledge, how often have you experienced the following situations as a result of sharing the datasets behind your published research with investigators outside your research groups? Include experiences from datasets shared outside your research groups through any mechanism, including public archiving, selected distribution, or shared individually upon request (for example, in response to an email request).
  a.  I formed new collaborations <!-- Field: Q9_1 / Q_4_1 -->
	b.  I formed new collaborations that led/are leading to publications <!-- Field: Q9_2 / Q_4_2 -->
	c.  I formed new collaborations that led/are leading to grants <!-- Field: Q9_3 / Q_4_3 -->
	d.  Others have used my datasets and formally cited me <!-- Field: Q9_4 / Q_4_4 -->
	e.  Others have used my datasets but did not formally cite me <!-- Field: Q9_5 / Q_4_5 -->
	f.  Others have found errors in my research through my datasets <!-- Field: Q9_7 / Q_4_6 -->
	g.  Others misinterpreted my datasets or used them inappropriately <!-- Field: Q9_8 / Q_4_7 -->
	h.  I have had ongoing research projects scooped by another scientist <!-- Field: Q9_9 / Q_4_8 -->
	i.  I have had future research plans scooped by another scientist <!-- Field: Q9_10 / Q_4_9 -->
	j.  The ability of a junior colleague in my group (graduate student, postdoctoral fellow, or junior faculty member) to publish subsequent research has been compromised <!-- Field: Q9_11 / Q_4_10 -->
	k.  My ability to receive commercial benefit from my research has been diminished <!-- Field: Q9_12 / Q_4_11 -->
	l.  I spent a lot of money preparing data for sharing <!-- Field: Q9_13 / Q_4_12 -->
	m.  I spent a lot of time preparing data for sharing <!-- Field: Q9_14 / Q_4_13 -->
	n.  I spent a lot of time answering questions from qualified researchers about my shared data <!-- Field: Q9_15 / Q_4_14 -->
	o.  Other (please specify) <!-- Field: Q9_16 - Q9_16_TEXT / Q_4_15 - Q_4_15_TEXT -->
  p.  I spent a lot of time answering questions from unqualified researchers about my shared data <!-- Field: Q9_17 / Q_4_16 -->

And again, this time limited to people with at least one paper with archived data:

And one more time, for people with no archived data:

And again, this time limited to people with at least one paper with archived data:

```
<!-- Field: Q11-->
5. To your knowledge, what are the policies of these stakeholders as they apply to online public archiving of the datasets associated with your (month) (year) publication in (journal)? Choose all that apply
	a.  The journal  <!-- Field: Q11_1 -->
	b.  Your funder(s)  <!-- Field: Q11_2 -->
	c.  Your employer or institution  <!-- Field: Q11_3 -->

<!-- Field: Q12 -->
6. Are all datasets associated with your (month) (year) publication in (journal) publicly available online?

Response options:
	a.  ALL datasets are publicly available online   <!-- Field: Q12_1 -->
	b.  SOME but not all of the datasets are publicly available online   <!-- Field: Q12_2 -->
	c.  NONE of the datasets are publicly available online   <!-- Field: Q12_3 -->
	d.  I don't know   <!-- Field: Q12_4 -->
	e.  We did not collect any data for that publication   <!-- Field: Q12_5 -->
	f. Other (please specify)   <!-- Field: Q12_6/Q12_6_TEXT -->

@@

p <- ggplot(tmp, aes(label_PolicyAdoptedAtPublication, class.Proportion*100, group=value, fill=value, alpha=value))+geom_bar(stat="identity", position="dodge")+xlab("JDAP Policy")+ylab("Proportion of responses (in %)")
p + theme_bw() + scale_fill_manual(values=c(rep("blue",2), rep("red",2), "grey", "green" )) + scale_alpha_manual(values=c(1,.5, 1,.5, 1,1))

@@

<!-- Field: Q13 -->
7. IF ALL or SOME datasets associated with this paper are publicly available, where are the datasets hosted?
	a.  Journal supplementary information   <!-- Field: Q13_1 -->
	b.  A repository for data or research output   <!-- Field: Q13_2 -->
	c.  A personal or lab website   <!-- Field: Q13_3 -->
	d.  The entire dataset is in the body of the paper   <!-- Field: Q13_4 -->
	e.  Other (please specify)   <!-- Field: Q13_5/Q13_5_TEXT -->
	f.  Not applicable   <!-- Field: Q13_6 -->

@@

<!-- Field: Q14 -->
8. If NONE or only SOME of the datasets are publicly available, which of the following reflect your experience for the dataset(s) not publicly available, for this paper? Choose all that apply.
	a.  The dataset has been submitted to a repository and is currently under embargo  <!-- Field: Q14_1 -->
	b.  We plan to publicly archive the dataset in the future  <!-- Field: Q14_2 -->
	c.  We plan to share the dataset upon request  <!-- Field: Q14_3 -->
	d.  We hadn't considered publicly archiving the dataset  <!-- Field: Q14_4 -->
	e.  We considered publicly archiving data and decided not to  <!-- Field: Q14_5 -->
	f.  We wanted to publicly archive the data but were not able to, due to resource limitations  <!-- Field: Q14_6 -->
	g.  We wanted to publicly archive the data but were not able to, for other reasons  <!-- Field: Q14_7 -->
	h.  Other (please specify)  <!-- Field: Q14_8/Q14_8_TEXT -->
	i.  Not applicable  <!-- Field: Q14_9 -->
	j.  We wanted to publicly archive the data but were not able to, due to licencing or funder restrictions  <!-- Field: Q14_10 -->

@@

levels <- c("The dataset has been submitted to a repository and is currently under embargo",
            "We plan to publicly archive the dataset in the future",
            "We plan to share the dataset upon request",
            "We hadn't considered publicly archiving the dataset",
            "We considered publicly archiving data and decided not to",
            "We wanted to publicly archive the data but were not able to, due to resource limitations",
            "We wanted to publicly archive the data but were not able to, for other reasons",           
            "We wanted to publicly archive the data but were not able to, due to licencing or funder restrictions",
            "Not applicable", "Other (please specify)")
cur_d$value <- ordered(cur_d$value, levels=levels)

@@

The questions below refer to the Joint Data Archiving Policy, an example policy statement adopted by ... <!-- TODO -->

<!-- Field: Q17 -->
9. How strongly do you disagree/agree with this statement? I would seek alternative journals if I learned that my target journal had adopted the Joint Data Archiving Policy (above).

<!--
Response options:
  i. Strongly Agree
  ii. Agree
  iii. Somewhat Agree
  iv. Neutral
  v. Somewhat Disagree
  vi. Disagree
  vii. Strongly Disagree
-->

>>>>

---
title: "Data Archiving Attitudes"
author: "Heather A. Piwowar, Todd J. Vision"
date: "March 20, 2018"
output:
  html_document:
    css: style.css
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
  md_document:
    variant: markdown_github
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
csl: peerj.csl
bibliography: input_references.bib
---

@@

# Abstract

# Introduction

@@

## The promise of data sharing

@@

## Data archiving journal policies

@@

A journal policy that requires authors to share the datasets that support their results as a condition of publication may impact the practices, attitudes, and opinions of data producing authors in many ways. The most direct impact is likely a change in frequency of data sharing, and a shift to data archiving locations specified in the policy.  Early studies suggest that authors who publish in journals with data sharing policies are more likely to share their data than authors who publish similar studies in similar journals `r citep(c(bib["piwowar2008a-revie"], bib["Piwowar_2011"]))`, but these investigations were correlative and did not attempt to directly measure the impact of journal policies over time.  Previous work emphasizes the importance of sharing data through best-practice archiving methods and locations.  Authors who share data upon request often discriminate against use and users `r citep(c(bib["campbell2008what-ar"], bib["reidpath2001data-sh"]))`.  The permanence of data sharing mechanisms is also a concern; email and lab-website based sharing choices have been shown to be impermanent solutions at best `r citep(c(bib["wren2008url-dec"], bib["wren2006e-mail"]))`.

@@

## Upcoming policy change is a useful opportunity for policy evaluation

Several journals plan to adopt the Joint Data Archiving Policy in early 2011, as described at http://datadryad.org/jdap.

@@

## Rigorous research is needed into the impact of journal policies

@@

## Research questions

@@

- How do authors’ attitudes, experiences, and practices around public data archiving change when the journals they publish in adopt mandatory data archiving policies?
- Are changes specific to the journals that implement the policies, or do they extend to other journals in the same subfield?
- Does publishing in a journal with a mandatory data archiving policy decrease an author's willingness to publish in a journal with a mandatory data archiving policy again?

Twelve other survey questions will allow secondary and exploratory analyses.

# Methods

Participants were recruited from corresponding authors who publish in specific journals between November 2010 and November 2013.  The responses of authors who published in JDAP-adopting journals were then compared, over time, to the responses of authors who published in similar journals that did not adopt JDAP.

The instrument used to collect data is be an online questionnaire. The questionnaire is designed to elicit the respondent’s practices, experiences, and opinions around public archiving of research datasets by asking about their opinion of community norms, recent behaviour with respect to a recently published paper, previous experiences with data sharing, and attitudes toward hypothetical journal policies

## Control jounrnals and sample size

The proposed study is subject to two different types of uncertainty:  uncertainty about the estimation of aggregate parameters, and also “uncertainty about the ability of the control group to reproduce the counterfactual of how the treated unit would have evolved in the absence of the treatment” (`r citep(bib["abadie2010synthet"])`. It was difficult to know how to conduct a formal power analysis in these conditions, so we  based our sample size decisions on the following information.

For populating the “treatment arm,” we asked all corresponding authors of all JDAP-adopting journals.

For populating the “control arms,” we recognized that the uncertainty is predominantly determined by the applicability of the controls.  Little information is currently available to determine, a priori, which journals will serve as appropriate controls.  As such, we decided that surveying twice as many control journals as treatment journals was an appropriate rule of thumb.  This balanced the need to avoid overpowering the study (and thereby increasing survey-fatigue and effort) with a desire to ensure the most useful results (anticipating that unexpected future events may render some controls inappropriate during the three years of this study, as well as recognizing that the quality of inference improves as the number of available comparison units increases `r citep(bib["abadie2010synthet"])`.

Control journals were selected by manual selection from a set of journals identified through ISI Journal Citation report categories as well as co-citation lists.  Attempts were made to choose journals with comparable impact factors and inclination to produce quantitative datasets.

Assuming a 20% response rate and a control arm that was twice as large as the treatment arm, we expected a maximum of 165 responses per month, or almost 6000 responses over the course of the three year study. The actual number of responses was expected to be lower than this because only invited each corresponding author the first time they published a paper.

Journal editors were informed of the survey before it began and given contact information in case they had questions, suggestions, or concerns.

## Participant recruitment

Participants were recruited from corresponding authors who published in specific journals between November 2010 and November 2013.  Specifically, email addresses were collected for papers published the previous month (or 3 previous months, for the initial month of recruitment) in each of the target journals through the ISI Web of Science.  No other identifying information (author names, article title, etc) was retained.

The email invitation included  a URL where subjects can learn more about the study and then read a fact sheet, consent, and participate if they choose.  This email contained contain an “opt out” option, which will remove their address from the reminder list.  Those who did not “opt out” were contacted by email again, a week later, politely asking them to take the survey if they haven’t already.

Recruitment email lists were compared across months, and emails only sent to those who have not previously received a recruitment email to reduce survey fatigue and ill-will.

The online survey was administered through Qualtrics, hosted by the University of North Carolina.  The online questionnaire took about 10 minutes to complete.

## Analysis

The responses of authors who published in JDAP-adopting journals were compared, over time, to  the responses of authors who published in similar journals that did not adopt JDAP.

@@

# Results

## Response rate

```{r, cache=TRUE, message=FALSE, echo=FALSE, warning=FALSE}

@@

[TODO: LIST THE JOURNALS IN CONTROL AND JDAP ARMS]

The invitations were sent out monthly between February 2011 and March 2014 to a total of $`r nrow(invitation_timing)`$ authors, and $`r nrow(d)`$ responded (corresponding to a response rate of $`r round(100*nrow(d)/nrow(invitation_timing))`$\%).

## Primary analysis

### Attitudes on data sharing and data archiving norms

The questions asked about the strength of agreement/disagreement with the following statements:

* It is the community norm in my field to share, with qualified researchers, datasets that support the results of peer-reviewed research articles. Include data sharing that occurs by any mechanism (e.g. emailing datasets upon request, posting datasets to lab websites, including datasets in journal supplementary information or data repositories).
* It is the community norm in my field to publicly archive all supporting datasets online, for use by anyone for any purpose, upon publication of a peer-reviewed research article.

@@

```{r, fig.width=10, fig.height=5, echo=FALSE, warning=FALSE, message=FALSE}

library(plyr)
library(dplyr)
library(ggplot2)
library(reshape2)
source("./src/plotLikert.R")

responses <- c("Strongly Agree","Agree","Somewhat Agree","Somewhat Disagree","Disagree","Strongly Disagree") # response categories from top to bottom
question_labels <- c('Q_1'='"Data sharing is a community norm."',
                     'Q_2'='"Data archiving is a community norm."')
tmp <- melt(d, measure.vars=c('Q_1','Q_2'))
tmp$variable <- as.character(tmp$variable)
tmp$variable <- ordered(nmap(tmp$variable, question_labels), levels=question_labels)
#tmp$variable <- gsub("Q_", "Question ", tmp$variable) # gsub("Q_", "Question ", tmp$variable)

p <- plot_likert(tmp, response='value', var_x='label_PolicyAdoptedAtPublication', var_facet="variable", response_order=responses, above_mid="Somewhat Agree", na.label="No response", xlab="% Responses", ylab="JDAP Policy")
print(p)
```

[TODO: REPHASE PARAGRAPHS BELOW]

```{r, message=FALSE, eval=TRUE, echo=FALSE, warning=FALSE}

@@

Similarily, the following table shows the estimated coefficients and 95\%-confidence intervals (transformed to odds ratios) for responses to *question 2*. According to the model estimates, the introduction of the JDAP data archiving policy significantly increases the odds of a positive response regarding *data archiving* by a factor of $1.57$ (CI=$[1.33; 1.90]$), while the remaining two predictors do not show any statistically significant effects.

@@

## Exploratory analysis

### Data sharing and archiving behavior

<!-- Field: Q12 -->
Are all datasets associated with your (month) (year) publication in (journal) publicly available online?

@@

@@

<!-- Field: Q13 -->
IF ALL or SOME datasets associated with this paper are publicly available, where are the datasets hosted?

@@

<!-- Field: Q14 -->
If NONE or only SOME of the datasets are publicly available, which of the following reflect your experience for the dataset(s) not publicly available, for this paper? Choose all that apply.

@@

levels <- c("The dataset has been submitted to a repository and is currently under embargo",
            "We plan to publicly archive the dataset in the future",
            "We plan to share the dataset upon request",
            "We hadn't considered publicly archiving the dataset",
            "We considered publicly archiving data and decided not to",
            "We wanted to publicly archive the data\nbut were not able to, due to resource limitations",
            "We wanted to publicly archive the data\nbut were not able to, for other reasons",           
            "We wanted to publicly archive the data\nbut were not able to, due to licencing or funder restrictions",
            "Not applicable", "Other (please specify)")
cur_d$value <- ordered(cur_d$value, levels=levels)

@@

<!-- Field: Q11-->
To your knowledge, what are the policies of these stakeholders as they apply to online public archiving of the datasets associated with your (month) (year) publication in (journal)? Choose all that apply

### Data archiving history

<!-- Field: Q7 -->
Do any of your published research papers have publicly archived datasets?

print(p)
```

### Experiences with consequences of data archiving

<!-- Field: Q9 -->
To your knowledge, how often have you experienced the following situations as a result of sharing the datasets behind your published research with investigators outside your research groups? Include experiences from datasets shared outside your research groups through any mechanism, including public archiving, selected distribution, or shared individually upon request (for example, in response to an email request).

 a.  I formed new collaborations <!-- Field: Q9_1 / Q_4_1 -->
	b.  I formed new collaborations that led/are leading to publications <!-- Field: Q9_2 / Q_4_2 -->
	c.  I formed new collaborations that led/are leading to grants <!-- Field: Q9_3 / Q_4_3 -->
	d.  Others have used my datasets and formally cited me <!-- Field: Q9_4 / Q_4_4 -->
	e.  Others have used my datasets but did not formally cite me <!-- Field: Q9_5 / Q_4_5 -->
	f.  Others have found errors in my research through my datasets <!-- Field: Q9_7 / Q_4_6 -->
	g.  Others misinterpreted my datasets or used them inappropriately <!-- Field: Q9_8 / Q_4_7 -->
	h.  I have had ongoing research projects scooped by another scientist <!-- Field: Q9_9 / Q_4_8 -->
	i.  I have had future research plans scooped by another scientist <!-- Field: Q9_10 / Q_4_9 -->
	j.  The ability of a junior colleague in my group (graduate student, postdoctoral fellow, or junior faculty member) to publish subsequent research has been compromised <!-- Field: Q9_11 / Q_4_10 -->
	k.  My ability to receive commercial benefit from my research has been diminished <!-- Field: Q9_12 / Q_4_11 -->
	l.  I spent a lot of money preparing data for sharing <!-- Field: Q9_13 / Q_4_12 -->
	m.  I spent a lot of time preparing data for sharing <!-- Field: Q9_14 / Q_4_13 -->
	n.  I spent a lot of time answering questions from qualified researchers about my shared data <!-- Field: Q9_15 / Q_4_14 -->
	o.  Other (please specify) <!-- Field: Q9_16 - Q9_16_TEXT / Q_4_15 - Q_4_15_TEXT -->
    p.  I spent a lot of time answering questions from unqualified researchers about my shared data <!-- Field: Q9_17 / Q_4_16 -->

Limited to people with at least one paper with archived data:

```{r, fig.width=10, fig.height=5, echo=FALSE, eval=FALSE}

# Everyone 
library(ggplot2)
library(reshape2)
source("./src/plotLikert.R")

```{r, fig.width=10, fig.height=5, echo=FALSE, eval=FALSE}

# For people with no archived data:

responses <- c( "Many times","A few times","Once","Never") # response categories from top to bottom
tmp <- melt(d_with_no_data, measure.vars=paste0('Q_4_',1:16 ))
tmp$variable <- letters[gsub("Q_", "", tmp$variable) %>% gsub("4_", "", .) %>% as.integer]
p <- plot_likert(tmp, response='value', var_x='label_PolicyAdoptedAtPublication_space', var_facet="variable", response_order=responses, above_mid="Once", na.label="No response", xlab="% Responses", ylab="JDAP Policy", include.N=FALSE)
p <- p + theme(axis.text.x=element_text(angle=-90, size=8))

### Attitudes on data reuse

<!-- Field: Q21 -->

responses <- c("Strongly Agree","Agree","Somewhat Agree","Somewhat Disagree","Disagree","Strongly Disagree") # response categories from top to bottom
